---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data


```{r}
library(dplyr)
dat<-read.csv("activity.csv")
dat<-mutate(dat,steps=as.numeric(steps),date=as.Date(date),interval=as.numeric(interval))

```




## What is mean total number of steps taken per day?
Here is histogram of total number of steps taken each day. I removed all the NA values for this part.
```{r}
library(dplyr)
dats<-dat
cc<-group_by(dats,by = date )
sumdats<-summarize(cc,sumsteps=sum(steps,na.rm=TRUE))
hist(sumdats$sumsteps,xlab= "total steps in each day",main="total number of steps taken each day")
meansteps=mean(sumdats$sumsteps)
print(meansteps) 
mediansteps=median(sumdats$sumsteps)
print(mediansteps)
```

The mean is `r meansteps`, and the median is `r mediansteps`.

## What is the average daily activity pattern?
I removed all the NA values for this part.
```{r}
library(dplyr)
dats<-dat
cc1<-group_by(dats,by = interval )
avedats<-summarize(cc1,ave=mean(steps,,na.rm=TRUE))
interval<-levels(factor(cc1$by))
# plot the time seris plot
plot(interval,avedats$ave,xlab= "time interval",ylab="average steps",type="l")
# maximum value
idx<-which.max(avedats$ave)
maxsteps<-interval[idx]
print(maxsteps)
```
`r maxsteps` is 5-minute interval with maximum number of average step

## Imputing missing values
I used the mean for that 5-minute interval for missing value. 
```{r}
missing<-which(is.na(dat$steps))
number_missing<-length(missing)
print(number_missing)
```
`r number_missing` are missing.  
making new dataset:
```{r}

new_dat<-dat
for (i in missing ) {new_dat$steps[i]<-avedats$ave[which(interval==dat$interval[i])]}

```
Histogram,mean, and median of new dataset

```{r}
library(dplyr)

cc2<-group_by(new_dat,by = date )
sumdats2<-summarize(cc2,sumsteps2=sum(steps))
hist(sumdats2$sumsteps2,xlab= "total steps in each day",main="total number of steps taken each day")
meansteps2=mean(sumdats2$sumsteps2)
print(meansteps2) 
mediansteps2=median(sumdats2$sumsteps2)
print(mediansteps2)
```
The mean of new dataset is `r meansteps2`, and the median is `r mediansteps2`.The mean and median are different because we are calculating them over fewer data.
## Are there differences in activity patterns between weekdays and weekends?

```{r}
new_dat<-mutate(new_dat,day= "weekday")
temp<-weekdays(new_dat$date)
new_dat$day[which(temp=="Saturday" |temp=="Sunday" ) ]<-"Weekend"


g1<-group_by(new_dat,interval,day)
ave1<-summarize(g1,ave=mean(steps))
interval<-levels(factor(g1$by))
# plot the time seris plot
library(dplyr)
library(lattice)
xyplot(ave ~ interval | day, data= ave1, type="l" ,layout= c(1,2),ylab ="Numbers of steps")
```

